---
title: "migration-shiny-prep"
author: "JJayes"
date: "06/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
theme_set(theme_light())
```


# Migration shiny prep
## Purpose

I want to create a shiny app that allows you to select a year from the census and see who migrated somewhere within sweden.

I have data from the IPUMS International dataset on Sweden.


### Reading in IPUMS data
```{r}
# library(ipumsr)
# library(here)
# 
# dat <- read_ipums_micro(
# 
#     ddi = here("migration-shiny-idea", "data", "ipumsi_00003.xml"),
#     data_file = here("migration-shiny-idea", "data", "ipumsi_00003.dat.gz")
# 
# )
# 
# dat %>% head() %>% view()
# 
# dat %>% colnames()
```

### Working with just the necessary info to start with

```{r}
# df <- dat %>% 
#     select(YEAR, PARSE, BPLSE2, NAMEFRST, SEX)
# 
# df %>% 
#     sample_n(100)
# 
# df %>% 
#     write_rds(here("migration-shiny-idea", "data", "migration_info.rds"), compress = "gz")
```

### Now working with a bit more info

I want to get information on the hisco codes, maybe gender also?

I'll stat with a small subset, maybe 1000 rows, just to get going.

PARSE Parish
BPLSE2 Parish of birth


FAMSIZE	Number of own family members in household

```{r}
# df <- dat %>%
#   sample_n(5000)

df %>% 
  select(YEAR, PARSE, BPLSE2, SEX, PERSONS, OCCISCO)

df %>% 
  count(OCCHISCO)

df %>% 
  count(OCCISCO)

df %>% 
  filter(OCCISCO != 99) %>% 
  count(YEAR, OCCISCO) %>% 
  complete(YEAR, OCCISCO, fill = list(n = 0)) %>% 
  ggplot(aes(YEAR, n, fill = factor(OCCISCO))) +
  geom_area(position = "fill")
```

Might want to know household type composition and hisco composition? And how this changes over time?

```{r}
df %>% 
  count(YEAR, SEX) %>% 
  ggplot(aes(YEAR, n, fill = factor(SEX))) +
  geom_area(position = "fill")
```

#### What about a ridge plot of household size over time?

```{r}
df %>%
  filter(PERSONS < 100) %>% 
  count(YEAR, PERSONS) %>% 
  # complete(YEAR, PERSONS, fill = list(n = 0)) %>% 
  ggplot(aes(PERSONS, fill = factor(YEAR))) +
  geom_density() +
  facet_wrap(~ YEAR)
  
library(ggridges)

df %>%
  filter(PERSONS < 60) %>% 
  count(YEAR, PERSONS) %>% 
  ggplot(aes(x = PERSONS, y = YEAR, fill = factor(YEAR))) +
  geom_density_ridges() +
  labs(x = "Household size",
       y = NULL) +
  theme(legend.position = "none")

df %>%
  filter(PERSONS < 60) %>% 
  mutate(household_type = if_else(HHTYPE < 7, "Family households", "Composite households")) %>% 
  count(YEAR, PERSONS, household_type) %>% 
  ggplot(aes(x = PERSONS, y = YEAR, 
             height = stat(density),
             fill = factor(YEAR))) +
  # geom_density_ridges(stat = "binline", bins = 70, scale = 0.45, draw_baseline = FALSE) +
  geom_density_ridges() +
  facet_wrap(~ household_type, nrow = 2) +
  labs(x = "Household size",
       y = NULL) +
  theme_ridges() +
  theme(legend.position = "none") 
```

Household types?

```{r}
df %>%
  count(YEAR, HHTYPE) %>% 
  ggplot(aes(YEAR, n, fill = factor(HHTYPE))) +
  geom_area(position = "fill")

df %>% count(HHTYPE)

HHTYPE_codes <- tibble(
  
  codes = c(1, 2, 3, 4, 6, 7, 8, 11, 99),
  descriptions = c(
    "One-person household",
    "Married/cohab couple, no children",
    "Married/cohab couple with children",
    "Single-parent family",
    "Extended family, relatives only",
    "Composite household, family and non-relatives",
    "Non-family household",
    "Group quarters",
    "Unclassifiable"
  )
  
)
```

NCOUPLES

```{r}
df %>% 
  count(NCOUPLES, YEAR) %>% 
  ggplot(aes(YEAR, n, fill = factor(NCOUPLES))) +
  geom_area(position = "fill")

df %>% 
  count(YEAR, SEX) %>% 
  ggplot(aes(YEAR, n, colour = factor(SEX))) +
  geom_line()

df %>% colnames()

df %>% count(YEAR, LABFORCE, SEX) %>% 
  ggplot(aes(YEAR, n, colour = factor(SEX))) +
  geom_line() +
  facet_wrap(~ LABFORCE)
```


```{r}
df %>% 
  count(SEX, OCCISCO) %>% 
  ggplot(aes(SEX, n, fill = factor(OCCISCO))) +
  geom_col(position = "fill")


df %>% 
  count(OCCISCO)

OCCISCO_codes <- tibble(
  
  codes = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 99),
  descriptions = c(
    "Legislators, senior officials and managers",
    "Professionals",
    "Technicians and associate professionals",
    "Clerks",
    "Service workers and shop and market sales",
    "Skilled agricultural and fishery workers",
    "Crafts and related trades workers",
    "Plant and machine operators and assemblers",
    "Elementary occupations",
    "Armed forces",
    "NIU (not in universe)"
  )
)
```


Mutations

```{r}
df %>% 
  mutate(SEX = factor(case_when(
    SEX == 1 ~ "Male",
    SEX == 2 ~ "Female"
  )))

```


```{r}
df <- read_rds(here("migration-shiny-idea", "data", "migration_info.rds"))

df %>% 
    count(PARSE, sort = T) 

df %>% 
    mutate(parish = parse_number(as.character(PARSE))) %>% 
    count(parish)
```


```{r}
map_file <- here("migration-shiny-idea", "maps", "Kommun_Sweref99TM_region.shp")

library(sf)
library(ggiraph)

map <- st_read(map_file)

# map %>% write_rds(here("migration-shiny-idea", "kommun_map.rds"))

instalmap %>%
  as_tibble() %>%
  count(KnKod)
```


```{r}
gg_iraph_map <- map %>% 
    ggplot() +
    geom_sf_interactive(aes(tooltip = KnNamn)) +
    theme_void()

girafe(
  ggobj = gg_iraph_map,
  width_svg = 8,
  height_svg = 4,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  )
)

```

### What about the parishes?

```{r}
library(leaflet)

map_file <- here("migration-shiny-idea", "maps", "forsamlingar_2021-01-01.shp")
map <- st_read(map_file)

map_wgs84 <- st_transform(map, "+init=epsg:4326")


map_leaflet <- map_wgs84 %>% 
    leaflet() %>% 
    addTiles() %>% 
    addPolygons(label=~namn)


```

### Fixing parish names

```{r}
df <- df %>% 
  # sample_n(10) %>% 
  mutate(kommun_living = str_sub(PARSE,1,nchar(PARSE)-5),
         kommun_living = case_when(
           
           nchar(kommun_living) == 3 ~ str_c("0", kommun_living),
           TRUE ~ kommun_living))

df %>% 
  count(kommun_living)
```

Great! There are just 288 kommuns that individuals are living in now, and we can match these to the list of kommun names that we can link to the map.

Then I can make a matrix that has the number of people born in a kommun on one axis and the number living in it on the other. Can use this axis to support the map.

```{r}
df <- df %>% 
  # sample_n(10) %>% 
  mutate(kommun_born = str_sub(BPLSE2,1,nchar(BPLSE2)-5),
         kommun_born = case_when(
           
           nchar(kommun_born) == 3 ~ str_c("0", kommun_born),
           TRUE ~ kommun_born))
```


```{r}
df_counts <- df %>% 
  count(YEAR, kommun_living, kommun_born)

# df_counts %>% 
#   write_rds(here("migration-shiny-idea", "data", "migration-counts.rds"), compress = "gz")
```


```{r}
library(readxl)
kommun_names <- read_excel(here("migration-shiny-idea", "data", "list-of-municipalities.xlsx")) %>% 
  janitor::clean_names()

kommun_names <- kommun_names %>% 
  filter(nchar(code) == 4)
```

Matching

```{r}
df_counts_named <- df_counts %>% 
  inner_join(kommun_names, by = c("kommun_living" = "code")) %>% 
  rename(kommun_living_name = name) %>% 
  inner_join(kommun_names, by = c("kommun_born" = "code")) %>% 
  rename(kommun_born_name = name)

# df_counts_named %>% 
#   write_rds(here("migration-shiny-idea", "data", "migration-counts-named.rds"), compress = "gz")

```


## Idea for flexdashboard

I want to have two plots in a single panel, and a table.

You choose a kommun and you see on a map how many people were born in it, and a plot with that same info on a column chart for, say, the 12 most prevalent municipalities.

```{r}
df_counts_named %>% 
  filter(YEAR == 1880,
         kommun_living == "0114") %>% 
  arrange(desc(n))
```

The function will look something like this:

You'll have a filter for the year, then you'll filter for kommun name, you'll join by birth to the map, and then draw the map. Not sure if I want to show shares or absolute numbers? Could do share on the map and aboslute numbers on the chart for the largest 5, for instance.
